---
title: "Assignment 4"
author: "By: Laurel Wee"
date: "11/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(effsize)
library(broom)
library(palmerpenguins)
library(ggbeeswarm)
library(kableExtra)
library(here)
library(lubridate)
library(dplyr)
```

### A. Introduction

### B. Data and analysis

### C. Exploratory findings

## i. Annual juvenile hare trap counts

```{r}
bonanza_hares <- read.csv(here("data","bonanza_hares.csv"))

bonanza_hares$date_formatted <- 
  as.Date(bonanza_hares$date, "%m/%d/%y") 
# Add a date_formatted column to get YYYY-MM-DD

hares_juvenile<- bonanza_hares %>% 
  filter(age == "j") %>%  
  # Create subset that is for juvenile hares
  mutate(juvenile_year = year(date_formatted)) %>%  
# add column that has the 'year'
  rename("site"= grid) %>% 
  mutate(sex = case_when(
    sex %in% "f" ~ "Female",
    sex %in% "m" ~ "Male",
    sex %in% "NA" ~ "Sex unknown")
  ) %>% 
  mutate(site_full = case_when(
    site %in% "bonbs" ~ "Bonanza Black Spruce",
    site %in% "bonmat" ~ "Bonanza Mature",
    site %in% "bonrip" ~ "Bonanza Riparian"
  ))

annual_hares_juvenile <- hares_juvenile %>% 
  count(juvenile_year) %>% 
  rename("sample_size"= n)

ggplot(data= hares_juvenile, 
       aes(juvenile_year))+
  geom_histogram(color= "black", 
                 bins = 30, binwidth = 1, alpha = .75)+
  labs(x = "Year", 
       y = "Number of Hares", 
       title = "Number of Juvenile Hares Trapped from 1999-2012 (Wee)")+
  theme_minimal()
```
# Figure 1: Number of juvenile snowshoe hares trapped each year between 1999-2012. 2002 and 2009 have no recorded juveniles trapped. Data: Kielland et. al. (2017).

# Table 1: Descriptive statistics (mean, median, min, and max) showing the distribution of juvenile showshoe hares caught annually. Data: Kielland et. al. (2017).

```{r}
juvenile_summary <- annual_hares_juvenile %>% 
  summarize(
    mean = mean(sample_size), 
    median  = median(sample_size),
    min  = min(sample_size),
    max = max(sample_size)) 

juvenile_summary %>% 
  kable(col.names = c("Mean Trapped", "Median Trapped" , "Minimum Trapped", "Maximum Trapped")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

## Frequency of snowshoe hares trapped annually

Summary stuff: No juvenile snowshoe hares were trapped in 2002 and 2009. The minimum and maximum juvenile hares trapped are 2 and 126 with the mean and median are 32 and 19.
Notable stuff: There is a negative trend in juvenile snowshoe hares trapped over time since 1999. 
Moving forward: Collect every year for the same number of days and traps...


### ii. Visualize juvenile hare weights

```{r}
juvenile_weight <- hares_juvenile %>% 
  select(site, sex, weight, age)

juvenile_weight_summary <- juvenile_weight %>% 
  group_by(sex) %>% 
  summarize(
    mean_weight = mean(weight, na.rm = T),
    median_weight = median(weight, na.rm = T), 
    sd_weight = sd(weight, na.rm = T),
    sample_size = n())
```


```{r}
ggplot() +
  geom_beeswarm(data = juvenile_weight, aes(x = sex, y= weight, color= sex))+
  scale_color_manual(values = c("goldenrod2","cyan4"), na.value = "coral") +
  geom_boxplot(data = juvenile_weight, aes(x = sex, y= weight), fill = NA, width = 0.2) +
  geom_point(data = juvenile_weight_summary, aes(x = sex, y = mean_weight), 
             color = "black",
             size = 3) +
  theme_minimal() +
  labs(x = "Juvenile Hare Sex",
       y = "Weight", title = " Weight of Juvenile Snowshoe Hares (Wee)")

```

Figure 2: 

Table 2: 
```{r}
juvenile_weight_summary %>% 
  kable(col.names = c("Sex", "Mean Weight (g)", "Median Weight (g)" , "Standard deviation (g)" , "Sample size")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

#The actual difference in means
#Effect size (Cohen's *d*)
#Outcome of a two-sample t-test

```

## iv. Relationship between juvenile weight and hind foot length.

```{r}
juvenile_weight_foot <- hares_juvenile %>% 
  select(age, sex, site_full,hindft)
```
### d. Summary

### e. Citations

